---
title: "SARGE Real Time Data Visualization Code"
author: "Stephanie J. Wilson"
date: Sys.Date()
output: pdf_document
---

#This code utilizes example data from Wilson et al., (in prep) that was collected at the COMPASS synoptic and CB-NERR site, The Goodwin Islands in Yorktown Virginia, USA. 

```{r setup, include=FALSE}

# Load necessary packages 
library(data.table); library(tidyverse)
library(readxl); library(lubridate)
library(ggpubr); library(cowplot)
library(plotrix)

# 

```

## Would be nice to have a table that is output that has the logger power in and out
#then to plot those data to look for battery level loss. 

## EXO2 Sonde (Xylem Analytics) Current Data 
```{r, echo=FALSE}

#Read in current water quality sonde (EXO2) data from SARGE
exo <- read.table("Example_Data/WILSON_ExoTable.dat", header=TRUE, skip=3, sep= ",")

#Take a quick look at the data 
head(exo)

#could consider putting these into a csv that then I read in for the EXO to be cleaner... 
colnames(exo) <- c("TIMESTAMP","RECORD","Date","Time","Chlorophyll_RFU","Chlorophyll_ugL",
              "Conductivity","FDOM_QSU","FDOM_RFU","NLF_conductivity",
              "ODO_sat","ODO_local", "ODO_MgL", "Pressure_psia", "Salinity_PPT", 
              "Specific_Conductivity_uScm", "BGA_PE_RFU", "BGA_PE_ugL",
              "TDS_mg_L","Turbidity_FNU","Wiper_Position_mv","pH","pH_mv","Temp_C", 
              "Depth_m","Battery_v","Cable_v","Wiper_Current_ma","sn","snn")


#Make a new timestamp column that is recognized as date + time measured in the timezone needed:
exo$TIMESTAMP <- ymd_hms(exo$TIMESTAMP, tz = "EST")

#Ensure that columns that should be numbers are numeric for plotting:
exo$Depth_m <- as.numeric(exo$Depth_m)
exo$Conductivity <- as.numeric(exo$Conductivity)


## Plot Waterlevel (Depth_m)
waterlev <- ggplot(exo, aes(x=TIMESTAMP, y=Depth_m)) + geom_point(color="blue") + ylim(-0.5,2.5) + xlim(c(as.POSIXct('2024-09-19 00:00:00', format = "%Y-%m-%d %H:%M:%S"),
     as.POSIXct('2024-09-28 00:00:00', format = "%Y-%m-%d %H:%M:%S"))) + theme_classic()
waterlev

## Plot conductivity 
cond <- ggplot(exo, aes(x=TIMESTAMP, y=Salinity_PPT)) + geom_point(color="purple") + 
   xlim(c(as.POSIXct('2024-09-19 00:00:00', format = "%Y-%m-%d %H:%M:%S"),
     as.POSIXct('2024-09-28 00:00:00', format = "%Y-%m-%d %H:%M:%S"))) + theme_classic()
cond

## Can plot as many or as few variables as needed, consider plotting what is useful for checks

```

## SontekIQ series (Xylem Analytics) Current Data 
```{r, echo=FALSE}

#read in hydrologic data from Sontek IQ (Xylem Analytics)
sontek <- read.table("Example_Data/WILSON_SontekTable.dat", header=TRUE, skip=1, sep= ",")
sontek <- sontek[-(1:3), ]
sontek <- as.tibble(sontek)

#Take a quick look at the tibble created
head(sontek)

#Make a new timestamp column that is recognized as date + time measured in the timezone needed:
sontek$TIMESTAMP <- ymd_hms(sontek$TIMESTAMP, tz = "EST")

#Ensure that columns that should be numbers are numeric for plotting:
sontek$Flowrate <- as.numeric(sontek$Flowrate)
sontek$Water_depth <- as.numeric(sontek$Water_depth)
sontek$Index_velocity <- as.numeric(sontek$Index_velocity)

## Plot Sontek Flow Data: 
flow <- ggplot(sontek, aes(x=TIMESTAMP, y=Flowrate)) + geom_point(color="black") +       xlim(c(as.POSIXct('2024-09-19 00:00:00', format = "%Y-%m-%d %H:%M:%S"),
     as.POSIXct('2024-09-28 00:00:00', format = "%Y-%m-%d %H:%M:%S"))) + theme_classic()
flow

## Plot Index Velocity or Water flow Velocity 
flow1 <- ggplot(sontek, aes(x=TIMESTAMP, y=Index_velocity)) + geom_point(color="black") +       
  xlim(c(as.POSIXct('2024-09-19 00:00:00', format = "%Y-%m-%d %H:%M:%S"),
     as.POSIXct('2024-09-28 00:00:00', format = "%Y-%m-%d %H:%M:%S"))) + theme_classic()
flow1

## Plot water depth measured by the sontekIQ 
sdepth <- ggplot(sontek, aes(x=TIMESTAMP, y=Water_depth)) + geom_point(color="darkblue") +
    xlim(c(as.POSIXct('2024-09-19 00:00:00', format = "%Y-%m-%d %H:%M:%S"),
     as.POSIXct('2024-09-28 00:00:00', format = "%Y-%m-%d %H:%M:%S"))) + theme_classic()
sdepth


#can plot other variables as needed 
```

## LI7810 CH4/CO2/H2O Trace Gas Analyzer (LI-COR) Current Data
```{r, echo=FALSE}
#Note that similar code could be used with a LI7820 N2O Analyzer 

#read in gas concentration data from the LI7810 (LI-COR)
flux <- read.table("Example_Data/WILSON_FLUX_7810.dat", header=TRUE, skip=1, sep= ",")
flux <- flux[-(1:2), ]

#Take a quick look at the dataframe created
head(flux)

#Make a new timestamp column that is recognized as date + time measured in the timezone needed:
flux$TIMESTAMP1 <- ymd_hms(flux$TIMESTAMP, tz = "EST")

#Ensure that columns that should be numbers are numeric for plotting:
flux$CH4d_ppm <- as.numeric(flux$CH4)
flux$CO2d_ppm <- as.numeric(flux$CO2)


## Plot CH4 Data (Note: these are raw concentrations, not in water concentrations)
ghg_CH4 <- ggplot(flux, aes(x=TIMESTAMP1, y=CH4d_ppm)) + geom_point(color="darkviolet")+
    xlim(c(as.POSIXct('2024-09-19 00:00:00', format = "%Y-%m-%d %H:%M:%S"),
    as.POSIXct('2024-09-28 00:00:00', format = "%Y-%m-%d %H:%M:%S"))) + ylim(0,500000)+
    theme_classic()
ghg_CH4

## Plot CO2 Data (Note: these are raw concentrations, not in water concentrations)
ghg_CO2 <- ggplot(flux, aes(x=TIMESTAMP1, y=CO2d_ppm)) + geom_point(color="darkred")+
    xlim(c(as.POSIXct('2024-09-19 00:00:00', format = "%Y-%m-%d %H:%M:%S"),
    as.POSIXct('2024-09-28 00:00:00', format = "%Y-%m-%d %H:%M:%S"))) + ylim(0,10000) + 
    theme_classic()
ghg_CO2


```

## RAD7 Radon Detector (Durridge) Current Data  
```{r, echo=FALSE}
#Note this code could be adapted for use with the RAD8 (Durridge)

#read in radon concentration data from the RAD7 (Durridge)
rad <- read.table("Example_Data/WILSON_Rad7Table.dat", header=TRUE, skip=1, sep= ",")
rad <- rad[-(1:2), ]

#Take a quick look at the dataframe created
head(rad)

#Make a new timestamp column that is recognized as date + time measured in the timezone needed:
rad$TIMESTAMP <- ymd_hms(rad$TIMESTAMP, tz = "EST")

#Ensure that columns that should be numbers are numeric for plotting:
rad$Radon_concentration_Rad7 <- as.numeric(rad$Radon_concentration_Rad7)
rad$RH_sample_Rad7 <- as.numeric(rad$RH_sample_Rad7)


## Plot Radon Data
radon <- ggplot(rad, aes(x=TIMESTAMP, y=Radon_concentration_Rad7)) + 
  geom_line(color="darkgreen", linewidth=1) +
    xlim(c(as.POSIXct('2024-09-19 00:00:00', format = "%Y-%m-%d %H:%M:%S"),
     as.POSIXct('2024-09-28 00:00:00', format = "%Y-%m-%d %H:%M:%S"))) + theme_classic() +  
  ylim(0,400)
radon

## Plot relative humidity data to determine desiccant status
RH  <- ggplot(rad, aes(x=TIMESTAMP, y=RH_sample_Rad7)) + geom_point(color="red") +
  xlim(c(as.POSIXct('2024-09-19 00:00:00', format = "%Y-%m-%d %H:%M:%S"),
     as.POSIXct('2024-09-28 00:00:00', format = "%Y-%m-%d %H:%M:%S"))) + theme_classic()
RH


```

## Plot variables together to look for patterns 
```{r, echo=FALSE}

#Plot Conductivity, Water Velocity, and Water Depth 
Water_Quality <- ggplot() + 
    geom_point(data=exo, aes(x=TIMESTAMP, y=Conductivity, color="Conductivity"), size=0.75) + 
    geom_point(data=sontek, aes(x=TIMESTAMP, y=Index_velocity*5, color="Flow Rate"), size=0.75) + 
    geom_point(data=sontek, aes(x=TIMESTAMP, y=Water_depth*5, color="Water Depth"), size=0.75) + 
    scale_color_manual(values = c("#5F464B", "#6BAA75", "#2978A0"))+
    xlim(c(as.POSIXct('2024-09-19 00:00:00', format = "%Y-%m-%d %H:%M:%S"),
     as.POSIXct('2024-09-28 00:00:00', format = "%Y-%m-%d %H:%M:%S"))) +
    scale_y_continuous( 
    "Conductivity (uS/cm)", limits = c(-5,40),
    sec.axis = sec_axis(~ . /5, name = "Water Depth (m) & Flow Rate (m/s)")) + 
     #xlim(152, 163) + 
    theme_classic() + 
    theme(legend.position = "bottom") + 
    labs(color=NULL)

Water_Quality

#Plot the Greenhouse Gas Concentrations together 
GHGs <- ggplot() + 
    geom_line(data=flux, aes(x=TIMESTAMP1, y=CO2d_ppm, color="CO2"), size=0.75) + 
    geom_line(data=flux, aes(x=TIMESTAMP1, y=CH4d_ppm/10, color="CH4"), size=0.75) + 
    scale_color_manual(values = c( "darkviolet", "darkred", "black"))+
    xlim(c(as.POSIXct('2024-09-19 00:00:00', format = "%Y-%m-%d %H:%M:%S"),
     as.POSIXct('2024-09-28 00:00:00', format = "%Y-%m-%d %H:%M:%S"))) +
    scale_y_continuous(
    "CO2 (ppm)", limits = c(0,40000),
    sec.axis = sec_axis(~ . *10, name = "CH4 (ppb)")) + 
   # xlim(152, 163) + 
    theme_classic() + 
    theme(legend.position = "bottom") + 
    labs(color=NULL)
GHGs

#Plot Radon and CH4 together 
Radon_ghg <- ggplot() + 
    geom_line(data=flux, aes(x=TIMESTAMP1, y=CH4d_ppm/100, color="CH4"), size=0.75) + 
    #geom_line(data=flux, aes(x=TIMESTAMP1, y=CO2d_ppm, color="CO2"), size=0.75) +
    geom_line(data=rad, aes(x=TIMESTAMP, y=Radon_concentration_Rad7, color="Radon"), size=0.75) + 
    #geom_line(data=exo, aes(x=fDOY, y=Depth_m*500), color="red") +
    scale_color_manual(values = c("darkviolet", "darkred", "darkgreen"))+
     xlim(c(as.POSIXct('2024-09-19 00:00:00', format = "%Y-%m-%d %H:%M:%S"),
     as.POSIXct('2024-09-28 00:00:00', format = "%Y-%m-%d %H:%M:%S"))) +
    scale_y_continuous(
    "Radon (Bq/m3)", limits = c(0,4000) ,
    sec.axis = sec_axis(~ . *100, name = "CH4 (ppm)")) + 
    theme_classic() + 
    labs(color=NULL)
Radon_ghg

ggarrange(Water_Quality, GHGs, ncol=2, nrow=1)


```

## Figure in SARGE Tech Note Paper:  
```{r, echo=FALSE}

#Plot Conductivity, Water Velocity, and Water Depth 
Water_Quality1 <- ggplot() + 
    geom_point(data=exo, aes(x=TIMESTAMP, y=Conductivity, color="Conductivity"), size=0.75) + 
    geom_point(data=sontek, aes(x=TIMESTAMP, y=Index_velocity*5, color="Flow Rate"), size=0.75) + 
    geom_point(data=sontek, aes(x=TIMESTAMP, y=Water_depth*5, color="Water Depth"), size=0.75) + 
    scale_color_manual(values = c("#5F464B", "#6BAA75", "#2978A0"))+
    xlim(c(as.POSIXct('2024-09-19 00:00:00', format = "%Y-%m-%d %H:%M:%S"),
     as.POSIXct('2024-09-28 00:00:00', format = "%Y-%m-%d %H:%M:%S"))) +
    scale_y_continuous( 
    "Conductivity (uS/cm)", limits = c(-5,40), #μ
    sec.axis = sec_axis(~ . /5, name = "Water Depth (m) & Flow Rate (m/s)")) + 
    labs(x=" ", title="A")+
    theme_classic() + 
    guides(color = guide_legend(override.aes = list(size = 5))) +
    theme(legend.position = "bottom", legend.text = element_text(size = 12)) + 
    labs(color=NULL)

Water_Quality1

#Plot the Greenhouse Gas Concentrations together 
GHGs1 <- ggplot() + 
    geom_point(data=flux, aes(x=TIMESTAMP1, y=CO2d_ppm, color="CO2"), size=0.75) + 
    geom_point(data=flux, aes(x=TIMESTAMP1, y=CH4d_ppm/10, color="CH4"), size=0.75) + 
    scale_color_manual(values = c( "darkviolet", "darkred"),
    labels = c("CO2" = expression(CO[2]), 
               "CH4" = expression(CH[4])) )+
    xlim(c(as.POSIXct('2024-09-19 00:00:00', format = "%Y-%m-%d %H:%M:%S"),
     as.POSIXct('2024-09-28 00:00:00', format = "%Y-%m-%d %H:%M:%S"))) +
    scale_y_continuous(
      name = expression("CO"[2] * " (ppm)"),
      limits = c(0, 40000),
      sec.axis = sec_axis(
        ~ . * 10,
      name = expression("CH"[4] * " (ppb)"))) + 
    labs(x=" ", title="B")+
    theme_classic() + 
    guides(color = guide_legend(override.aes = list(size = 5))) +
    theme(legend.position = "bottom", legend.text = element_text(size = 12)) + 
    labs(color=NULL)
GHGs1

figure <- ggarrange(Water_Quality1, GHGs1, ncol=2, nrow=1)
figure

```
